---
title: "Comparison_to_prior_studies"
output: html_document
date: "2024-08-13"
---

#####################################################
Reanalyze Lea et al. 2018 results using new pipelie
#####################################################


# count data can be found here: 


```{r}
library(tidyverse)
```


```{r}
eLife<- read.delim('Lea_eLife_mSTARR_counts.txt')
methylated_counts<-eLife[,c(14:23)]
```

#methylated replicates only
```{r}
type <-c("DNA", "DNA", "DNA", "DNA", "DNA", "RNA", "RNA", "RNA", "RNA", "RNA")
design1<-model.matrix(~type)
dge <- DGEList(counts=(methylated_counts))
dge1 <- calcNormFactors(dge)
```

```{r}
v_methonly <- voomWithQualityWeights(dge1,design=design1,plot=F)
```

```{r}
fit <-lmFit(v_methonly,design1)
fit_dna <- eBayes(fit)
se.coef <- sqrt(fit_dna$s2.post) * fit_dna$stdev.unscaled
fit_dna_df <- as.data.frame(cbind(fit_dna$p.value[,2],fit_dna$coefficient[,2],se.coef[,2]))
colnames(fit_dna_df)<-c('pval_meth','beta_meth','se_meth')
fit_dna_df$id<-1:dim(fit_dna_df)[1]
fit_dna_df$q_meth<-qvalue(fit_dna_df[,1])$qvalues
fit_dna_keep01<-subset(fit_dna_df,(q_meth<0.01 & beta_meth>0))
dim(fit_dna_keep01) #1031 regulatory windows
dim(fit_dna_df) #756814 total windows analyzed
nrow(fit_dna_keep01)/nrow(fit_dna_df) #0.14%
fit_dna_df$info<-'not_regulatory'
fit_dna_df$info[which(fit_dna_df$id %in% fit_dna_keep01$id)]<-'regulatory'
fit_dna_df$site<-eLife$site
write.table(fit_dna_df,'AJL_eLife_methonly.txt',row.names=T,sep='\t',quote=F)
```

#unmethylated replicates only
```{r}
unmethylated_counts<-eLife[,c(2:13)]
```

```{r}
type <-c("DNA", "DNA", "DNA", "DNA", "DNA", "DNA", "RNA", "RNA", "RNA", "RNA", "RNA", "RNA")
design1<-model.matrix(~type)
dge <- DGEList(counts=(unmethylated_counts))
dge1 <- calcNormFactors(dge)
```

```{r}
v_unmethonly <- voomWithQualityWeights(dge1,design=design1,plot=F)
```

```{r}
fit <-lmFit(v_unmethonly,design1)
fit_dna <- eBayes(fit)
se.coef <- sqrt(fit_dna$s2.post) * fit_dna$stdev.unscaled
fit_dna_df <- as.data.frame(cbind(fit_dna$p.value[,2],fit_dna$coefficient[,2],se.coef[,2]))
colnames(fit_dna_df)<-c('pval_unmeth','beta_unmeth','se_unmeth')
fit_dna_df$id<-1:dim(fit_dna_df)[1]
fit_dna_df$q_unmeth<-qvalue(fit_dna_df[,1])$qvalues
fit_dna_keep01<-subset(fit_dna_df,(q_unmeth<0.01 & beta_unmeth>0))
dim(fit_dna_keep01) #5891 regulatory windows
dim(fit_dna_df) #756814 total windows analyzed
nrow(fit_dna_keep01)/nrow(fit_dna_df) #0.78%
fit_dna_df$info<-'not_regulatory'
fit_dna_df$info[which(fit_dna_df$id %in% fit_dna_keep01$id)]<-'regulatory'
fit_dna_df$site<-eLife$site
write.table(fit_dna_df,'AJL_eLife_unmethonly.txt',row.names=T,sep='\t',quote=F)
```

#What percent of tested windows are regulatory?
```{r}
eLife_unmeth<-read.table('AJL_eLife_unmethonly.txt')
eLife_meth<-read.table('AJL_eLife_methonly.txt')

dim(eLife_meth) #756814
eLife_meth_reg<-subset(eLife_meth, info == "regulatory"); dim(eLife_meth_reg) #1031
#1031/756814

dim(eLife_unmeth) #756814
eLife_unmeth_reg<-subset(eLife_unmeth, info == "regulatory"); dim(eLife_unmeth_reg) #5891
#5891/756814
```


# Are results similar for windows measured in Lea et al. 2018 and the current study?

```{r}
MethModel<-read.delim('model1_methonly_results.txt')
UnmethModel <- read.delim('model1_unmethonly_results.txt')
```

```{r}
MethModsites <- rownames(MethModel)
MethModstartsites<- sub('^([^_]+_[^_]+).*', '\\1', MethModsites)
MethModel$site <- MethModstartsites

MethModel_eLife <- merge(MethModel, eLife_meth, by="site") %>%
  na.omit(); dim(MethModel_eLife) #211,837 sites tested for MD activity in both datasets

UnmethModsites <- rownames(UnmethModel)
UnmethModstartsites<- sub('^([^_]+_[^_]+).*', '\\1', UnmethModsites)
UnmethModel$site <- UnmethModstartsites

UnmethModel_eLife <- merge(UnmethModel, eLife_unmeth, by="site") %>%
  na.omit(); dim(UnmethModel_eLife) #211,837 sites tested for MD activity in both datasets
```

#fischers exact test to see if more likely to call enhancers on the same regions than would be expected by chance.

```{r}
fisher.test(table(MethModel_eLife$info.x, MethModel_eLife$info.y))
#odds ratio= 79.65 for methylated sites

fisher.test(table(UnmethModel_eLife$info.x, UnmethModel_eLife$info.y))
#odds ratio= 54.86 for unmethylated sites
```

#correlation in effect sizes for windows in common
```{r}
# all windows in common- methylated
cor.test(MethModel_eLife$beta_meth.x, MethModel_eLife$beta_meth.y) #R=0.18

# windows that are regulatory in current study - methylated
MethCommonSites_sig <- MethModel_eLife %>%
  filter(info.x == "regulatory") #700 of our regulatory windows are in Lea et al. 2018 dataset
cor.test(MethCommonSites_sig$beta_meth.x, MethCommonSites_sig$beta_meth.y) #R=0.57

# all windows in common- unmethylated
cor.test(UnmethModel_eLife$beta_unmeth.x, UnmethModel_eLife$beta_unmeth.y) #R=0.32 

# windows that are regulatory in current study - unmethylated
UnmethCommonSites_sig <- UnmethModel_eLife %>%
  filter(info.x == "regulatory") #2520 of our regulatory windows are in Lea et al. 2018 dataset
cor.test(UnmethCommonSites_sig$beta_unmeth.x, UnmethCommonSites_sig$beta_unmeth.y) #R=0.60
```



#########################################################
Reanalyze Johnston et al. 2024 results using new pipelie
#########################################################

count data can be found here:

```{r}
library(tidyverse)
```

```{r}
RAJ<- read.table('Johnston_eLife_mSTARR_counts_K562.txt', h=T)

methylated<-c("L31223", "L31224", "L31225" ,"L31232", "L31233", "L31234", "L31259", "L31260", "L31261", "L31268", "L31269","L31270")

unmethylated<-c("L31241", "L31242", "L31243","L31250", "L31251", "L31252", "L31277", "L31278", "L31279", "L31286", "L31287", "L31288")
```

#unmethylated replicates only
```{r}
unmethylated_counts<-(RAJ[,unmethylated])
```

```{r}
type <-c("RNA", "RNA", "RNA", "RNA", "RNA", "RNA", "DNA", "DNA", "DNA", "DNA", "DNA", "DNA")
design1<-model.matrix(~type)
dge <- DGEList(counts=unmethylated_counts)
dge1 <- calcNormFactors(dge)
```

```{r}
v_unmethonly <- voomWithQualityWeights(dge1,design=design1,plot=F)
```

```{r}
fit <-lmFit(v_unmethonly,design1)
fit_dna <- eBayes(fit)
se.coef <- sqrt(fit_dna$s2.post) * fit_dna$stdev.unscaled
fit_dna_df <- as.data.frame(cbind(fit_dna$p.value[,2],fit_dna$coefficient[,2],se.coef[,2]))
colnames(fit_dna_df)<-c('pval_unmeth','beta_unmeth','se_unmeth')
fit_dna_df$id<-1:dim(fit_dna_df)[1]
fit_dna_df$q_unmeth<-qvalue(fit_dna_df[,1])$qvalues
fit_dna_keep01<-subset(fit_dna_df,(q_unmeth<0.01 & beta_unmeth>0))
dim(fit_dna_keep01) #9648 regulatory windows
dim(fit_dna_df) #497815 total windows analyzed
nrow(fit_dna_keep01)/nrow(fit_dna_df) #1.9%
fit_dna_df$info<-'not_regulatory'
fit_dna_df$info[which(fit_dna_df$id %in% fit_dna_keep01$id)]<-'regulatory'
fit_dna_df$site<-RAJ$site
write.table(fit_dna_df,'RAJ_eLife_unmethonly.txt',row.names=T,sep='\t',quote=F)
```


#methylated replicates only
```{r}
methylated_counts<-(RAJ[,methylated])
```

```{r}
type <-c("RNA", "RNA", "RNA", "RNA", "RNA", "RNA", "DNA", "DNA", "DNA", "DNA", "DNA", "DNA")
design1<-model.matrix(~type)
dge <- DGEList(counts=methylated_counts)
dge1 <- calcNormFactors(dge)
```

```{r}
v_methonly <- voomWithQualityWeights(dge1,design=design1,plot=F)
```

```{r}
fit <-lmFit(v_methonly,design1)
fit_dna <- eBayes(fit)
se.coef <- sqrt(fit_dna$s2.post) * fit_dna$stdev.unscaled
fit_dna_df <- as.data.frame(cbind(fit_dna$p.value[,2],fit_dna$coefficient[,2],se.coef[,2]))
colnames(fit_dna_df)<-c('pval_meth','beta_meth','se_meth')
fit_dna_df$id<-1:dim(fit_dna_df)[1]
fit_dna_df$q_meth<-qvalue(fit_dna_df[,1])$qvalues
fit_dna_keep01<-subset(fit_dna_df,(q_meth<0.01 & beta_meth>0))
dim(fit_dna_keep01) #8533 regulatory windows
dim(fit_dna_df) #497815 total windows analyzed
nrow(fit_dna_keep01)/nrow(fit_dna_df) #1.7%
fit_dna_df$info<-'not_regulatory'
fit_dna_df$info[which(fit_dna_df$id %in% fit_dna_keep01$id)]<-'regulatory'
fit_dna_df$site<-RAJ$site
write.table(fit_dna_df,'RAJ_eLife_methonly.txt',row.names=T,sep='\t',quote=F)
```


#what percent of tested windows are regulatory?
```{r}
RAJ_meth<-read.table('RAJ_eLife_methonly.txt', h=T)
RAJ_unmeth<-read.table('RAJ_eLife_unmethonly.txt', h=T)

dim(RAJ_meth) #497,815 windows tested
RAJ_meth_reg<-subset(RAJ_meth, info == "regulatory") #8,533 regulatory. 
#8533/497815

dim(RAJ_unmeth) #497,815 windows tested
RAJ_unmeth_reg<-subset(RAJ_unmeth, info == "regulatory"); dim(RAJ_unmeth_reg) #9,648 regulatory. 

```


# are the results similar?
```{r}
MethModel<-read.delim('model1_methonly_results.txt')
UnmethModel <- read.delim('model1_unmethonly_results.txt')

MethModsites <- rownames(MethModel)
MethModstartsites<- sub('^([^_]+_[^_]+).*', '\\1', MethModsites)
MethModel$site <- MethModstartsites

MethCommonWindows <- merge(MethModel, RAJ_meth, by="site"); dim(MethCommonWindows)
#135,595 sites in common

UnmethModsites <- rownames(UnmethModel)
UnmethModstartsites<- sub('^([^_]+_[^_]+).*', '\\1', UnmethModsites)
UnmethModel$site <- UnmethModstartsites

UnmethCommonWindows <- merge(UnmethModel, RAJ_unmeth, by="site"); dim(UnmethCommonWindows)
#135,595 sites in common
```


#fischers exact test to see if more likely to call enhancers on the same regions than would be expected by chance.
```{r}
fisher.test(table(MethCommonWindows$info.x, MethCommonWindows$info.y)) #odds ratio 26.3

fisher.test(table(UnmethCommonWindows$info.x, UnmethCommonWindows$info.y)) #odds ratio 23.3
```

#correlation in effect sizes for windows in common
```{r}
# all windows- methylated
cor.test(MethCommonWindows$beta_meth.x, MethCommonWindows$beta_meth.y) #R=0.10 

# windows significant in current study- methylated
MethCommonWindows_sig <- MethCommonWindows %>%
  filter(info.x == "regulatory") #427 of our regulatory windows are in RAJ dataset

cor.test(MethCommonWindows_sig$beta_meth.x, MethCommonWindows_sig$beta_meth.y) #R= 0.48 


# all windows- unmethylated
cor.test(UnmethCommonWindows$beta_unmeth.x, UnmethCommonWindows$beta_unmeth.y) #R=0.10 << all windows

# windows significant in current study- unmethylated
UnmethCommonWindows_sig <- UnmethCommonWindows %>%
  filter(info.x == "regulatory") #Only 1,770 of our regulatory windows are in RAJ dataset

cor.test(UnmethCommonWindows_sig$beta_unmeth.x, UnmethCommonWindows_sig$beta_unmeth.y) #R= 0.58
```


